# Case Study Project - Ukranian Coffee Shops

The dataset which will be used contains information about coffee shops in Ukraine. I will determine the market segments in this market along with the key features so this can be presented to the marketing team. I will focus initially on data wrangling to get the dataset in shape for modeling, then will dive deep into principal component analysis to determine the key market segments.

## Objective 

The objective has been set and detailed in the document below:

https://s3.amazonaws.com/talent-assets.datacamp.com/Certification+Case+Study+-+Coffee+Shops+-+Prework.pdf

```{r, message=FALSE}
# Settings used

knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	fig.align='center'
)

```

## Load the tidyverse and dataset

```{r}
# Load the tidyverse and janitor libraries

library(tidyverse)
library(janitor)

# Load the dataset

df <- read_csv("coffee_shops.csv")

df_clean <- df %>%
            clean_names()

```

Place_name is irrelevant in the dataset so will be deleted

## Understanding our data

```{r}

# The code below will produce an interactive table using the datatable package

df_clean <- df_clean %>%
            select(-place_name) %>%
            mutate(region = as.factor(region)) %>%
            mutate(place_type = as.factor(place_type)) 

# Below code removes irrelevant data point related to an appliance repair service

# Dine in option describes whether there is (or isn't) a delivery option so will set missing values as FALSE 

# Take-out option describes whether there is (or isn't) a Take-out option so will set missing values as FALSE 

# Below code categorises price into Low, Medium, and High 

df_clean <- df_clean %>%
  filter(!place_type == "Appliance repair service") %>%
  mutate(dine_in_option = if_else(is.na(dine_in_option),FALSE,TRUE)) %>%
  mutate(takeout_option = if_else(is.na(takeout_option),FALSE,TRUE)) %>%
  mutate(delivery_option = if_else(dine_in_option == FALSE & takeout_option == FALSE, TRUE, delivery_option)) %>%
  mutate(price = if_else(price == "$", "Low", price)) %>%
  mutate(price = if_else(price == "$$", "Medium", price)) %>%
  mutate(price = if_else(price == "$$$", "High", price))



```


### Interactive map showing region details

```{r}

# The code below will produce an interactive table using the datatable package. I also use basic JavaScript and CSS to style the table headers.

library(DT)

datatable(df_clean, options = list(
  initComplete = JS("
    function(settings, json) {
      $(this.api().table().header()).css({
        'background-color': '#009879',
        'color': '#fff'
      });
    }")
))

```

### Price categories histogram

```{r}

# Plots a histogram of the different price categories   

df_clean %>%
  count(price) %>%
  ggplot(aes(price, n, fill = price)) +
  geom_col() + 
  ylab("Number within each price category")

```

Price category with missing values is big piece of the dataset.

### Ratings distribution per region

```{r}

# This will determine ratings distribution per region and arrange orderly by avg_rating so regions showing first are those with the lowest average rating

  df_clean %>%
  group_by(region) %>%
  mutate(avg_rating = mean(rating)) %>%
  ungroup() %>%
  mutate(region = fct_reorder(region, avg_rating)) %>%
  ggplot(aes(rating, fill = region)) +
  geom_histogram() + 
  facet_wrap(~region) + 
  theme(legend.position = "none")

```

Rating appears to be influenced by region but not a main factor. However, variability shown by the disparity in range shows that there are potential other features that influence market segmentation. The boxplot below shows this more clearly.

### Average rating per region

```{r}
# Below code will produce a box plot

  df_clean %>%
  ggplot(aes(rating, reorder(region, rating, na.rm = TRUE))) +
  geom_boxplot() + 
  ylab("region")

```


### Average rating by price category

```{r}

# Construct our plot using ggplot

  df_clean %>%
  group_by(price) %>%
  summarise(avg_rating = mean(rating, na.rm = TRUE)) %>%
  ggplot(aes(price, avg_rating, fill = price)) +
  geom_col() + 
   ylab("Average rating") + 
   xlab("Price category")

```

Clearly, there is no significant difference in rating between the different price 
categories. NA category (i.e. missing) is slightly higher but not significantly. 

### Average review by price category
 
```{r}

# Construct our plot using ggplot

  df_clean %>%
  group_by(price) %>%
  summarise(avg_reviews = mean(reviews, na.rm = TRUE)) %>%
  ggplot(aes(price, avg_reviews, fill = price)) +
  geom_col()

```

This reflects that there is a much bigger segment of people who visit medium priced restaurants. 

### Total reviews by price category

```{r}

# Construct using ggplot 

  df_clean %>%
  group_by(price) %>%
  summarise(total_reviews = sum(reviews, na.rm = TRUE)) %>%
  ggplot(aes(price, total_reviews, fill = price)) +
  geom_col() + 
  xlab("Price category") + 
  ylab("Total reviews")

```

Now I comfortable that there is no clear imputation of missing price categories so will remove from the dataset. 

### Total reviews by region

```{r}

# Visualise total reviews by region to help us better understand the dataset 

region_reviews <-  df_clean %>%
  group_by(region) %>%
  summarise(total_reviews = sum(reviews, na.rm = TRUE),
            avg_rating = mean(rating, na.rm = TRUE))

  df_clean %>%
  group_by(region) %>%
  summarise(total_reviews = sum(reviews, na.rm = TRUE)) %>%
  mutate(region = fct_reorder(region, total_reviews)) %>%
  ggplot(aes(region, total_reviews, fill = region)) +
  geom_col() + 
  coord_flip() + 
  xlab("Region") + 
  ylab("Total reviews")

```


### Leaflet map

```{r, fig.height=8, fig.width= 8, fig.align='center'}

# Below code leverages the leaflet library. To set regions as markers in our map, I have sought the longitudes and latitudes from the Geodatos website as provided in the reference below.

library(leaflet)

sf <- data.frame(lng = c(35.04066, 36.25272, 36.25272, 33.39404, 30.5238, 24.02324, 37.5413, 30.74383, 34.55367, 35.11714),
                 lat = c(48.46664, 49.98081, 46.63695, 47.90572, 50.45466, 49.83826, 47.09514, 46.48572, 49.58925, 47.85167),
                 total_reviews = region_reviews$total_reviews,
                 popup = region_reviews$region,
                 region = region_reviews$region,
                 rating = region_reviews$avg_rating,
                 stringsAsFactors = FALSE)

# Below will add colors to the different regions and contrast will be higher by regions that have higher overall total reviews.

sf <- sf %>% 
  mutate(region = fct_reorder(region, total_reviews))

pal <- colorFactor(
  palette = "YlOrRd",
  domain = sf$region
)

# This is the final code to construct the map using the lng, lat.

leaflet(sf) %>%
  addTiles() %>%
  addCircleMarkers(lng = ~lng, 
                   lat = ~lat,
                   weight = 20,
                   radius = ~ log(sf$total_reviews),
                   color = ~ pal(region),
                   label = paste(" Region = ", sf$region, ",",
                                 "Total Reviews = ",sf$total_reviews,",",
                                 "Average Rating = ", round(sf$rating,1))) %>%
    addLegend(position = "bottomleft", pal = pal, values = ~region)

```
 

### Total reviews by place type


```{r}

  df_clean %>%
  group_by(place_type) %>%
  summarise(total_reviews = sum(reviews, na.rm = TRUE)) %>%
  mutate(place_type = fct_reorder(place_type, total_reviews)) %>%
  ggplot(aes(place_type, total_reviews, fill = place_type)) +
  geom_col() + 
  coord_flip()

```

Most total reviews come from either coffee shop, cafe and espresso bar.

As Kraydieh insightfully answered on Quora when addressed a question on differences between an espresso bar and co"it’s a bit bigger than a kiosk, it has a bar, stools underneath, but it is much smaller than an a regular shop."

### Final data cleaning touches before modeling

```{r}

df_clean_place <-df_clean %>%
                  mutate(place_type = str_to_lower(place_type),
                         place_type = case_when(str_detect(place_type,"store") ~ "coffee store",
                                                str_detect(place_type,"coffee shop") ~ "coffee shop",
                                                str_detect(place_type,"espresso bar") ~ "espresso bar",
                                                str_detect(place_type,"cafe") ~ "cafe",
                                                TRUE ~ "other"))

df_parsed <- df_clean_place %>%
              mutate(delivery_option = as.integer(delivery_option),
                     dine_in_option = as.integer(dine_in_option),
                     takeout_option = as.integer(takeout_option)) %>%
              drop_na(price, delivery_option) %>%
              pivot_wider(names_from = price, values_from = rating, values_fill = 0)

library(DT)

datatable(df_parsed, options = list(
  initComplete = JS("
    function(settings, json) {
      $(this.api().table().header()).css({
        'background-color': '#009879',
        'color': '#fff'
      });
    }")
))


```


## Modeling using Principal Component Analysis

```{r}

# Load the tidy models library

library(tidymodels)

# Use reviews to set the recipe or formula.
# PCA will not use region and place_type
# Use log reviews due to the high number of reviews

pca_rec <- recipe(reviews ~ ., data = df_parsed) %>%
            update_role(region, place_type, new_role = "id") %>%
            step_log(reviews) %>%
            step_normalize(all_predictors()) %>%
            step_pca(all_predictors())

pca_prep <- prep(pca_rec)

pca_prep

```

Below shows that we have 2 ids, 1 outcome which is what we are interested in (reviews) to understand market share segments, and 6 predictors. 

```{r}

# Code below will tidy our data by showing contributing values by the computed features and classifies these into different components

tidied_pca <- tidy(pca_prep, 3)

# Now will plot the PCA terms (features) by each of the six components

tidied_pca %>%
  mutate(component = fct_inorder(component)) %>%
  ggplot(aes(value, terms, fill = terms)) + 
  geom_col(show.legend = FALSE) + 
  facet_wrap(~component) + 
  labs(y = NULL)

```


```{r}

# There are many features so eliminate irrelevant ones by showing the top 6 for each PCA component 

# Will also load the tidytext library to reorder within to produce a prettier chart

library(tidytext)

tidied_pca %>%
   filter(component %in% c("PC1", "PC2", "PC3", "PC4")) %>%
   group_by(component) %>%
   top_n(6, abs(value)) %>%
   mutate(terms = reorder_within(terms, abs(value), component)) %>%
   ggplot(aes(abs(value), terms, fill = value > 0)) + 
   geom_col() + 
   facet_wrap(~component, scales = "free_y") +
   scale_y_reordered() +
   labs(x = "Absolute value of contribution", y = NULL, fill = "Positive?")

```

#### Main takeaways from the above results

* The first principal component tells whether the price is medium or not (average cafe or not)

* The second PC tells us whether the coffee shop entertains either takeouts or deliveries.

* The third PC is whether the coffee shop is marketed as a niche as either low or high-end

* The fourth PC is whether the coffee shop does dine-in or delivery

## Significance testing

```{r}
# Finally, we will perform a linear model to test the most significant component
# Note we don't need to re-do a log of reviews as it is already built-in the PCA structure.

pca_lm <- juice(pca_prep) %>%
    select(-region, -place_type) %>%
    lm(reviews ~ ., data = .)

summary(pca_lm)
```


## Key Findings 

From the above we can tell there is a significant evidence that PC4 and PC1 are contributing factors to the overall number of reviews given.

Explaining this further, we could infer **4 main market segments** and these are : 

* Medium price category
* High/Low price category (niche market price)
* Diners or "Dine-ins"
* Delivery stores

**However, it would be interesting to test the significance of this assertion had we had data on the total number of visits.**

 
### References 
 
* https://www.quora.com/What-is-the-difference-between-a-cafe-coffee-shop-espresso-bar-and-other-establishments-that-specializes-in-serving-coffee-based-drinks

* https://www.geodatos.net/en/coordinates/ukraine
